---
title: "Synapse Demo - File Uploads"
author: "Aryton Tediarjo"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

This R markdown is used for showcasing basic interaction with Synapse using RStudio. It will cover on how to utilize our Synapse R client to do basic uploading/downloading files.

## Import Libraries
```{r warning=FALSE, message=FALSE}
library(data.table)
library(tidyverse)
library(synapser)
library(synapserutils)
synLogin()
```

## Global Variables
```{r}
SYNAPSE_PARENT_ID <- "syn25257185" # where you will be saving all your files
FILEPATH <- file.path("data", "sample_1.tsv")
```


## Part 1: Uploading to Synapse (1 File)

### Step 1. Read Arbitrary Data
```{r warning=FALSE}
data <- fread(FILEPATH) %>% 
    tibble::as_tibble(.)
data
```

### Step 2. Create File Entity Locally
```{r}
file_entity <- synapser::File(FILEPATH, parentId = SYNAPSE_PARENT_ID)
file_entity
```

### Step 3. Store to Synapse
```{r}
synStore(file_entity)
```

## Part 2: Files Versioning, Provenance, Annotations

File Entity Components:

Versioning: 
Let's say you want to add more data into the previously stored dataframe, Synapse will provide you the capability to version your files everytime you update it so that you can always access your previous data.

Provenance:
You can also track the sources of how the data is generated

Annotations:
Key-Value pair attached to a data

### a. Versioning
```{r}
data_new <- data %>% 
    add_row(tibble_row(
        user = "jane doe", 
        createdOn = lubridate::as_date(lubridate::now()),
        x = 4, y = 0))
data_new %>% 
    write.table(., FILEPATH, sep = "\t",
                row.names = FALSE, 
                quote = FALSE)
```

```{r echo=FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
new_file_entity <- synapser::File(FILEPATH, parentId = SYNAPSE_PARENT_ID)
store_to_synapse <- synStore(new_file_entity)
```

### b. Files Provenance
```{r}
USED_SYNAPSE_ID <- "syn23555388"
USED_NB_GIT_URL <- "https://github.com/arytontediarjo/synapse-demo/blob/main/file_upload.Rmd"

activity = synapser::Activity(
    "generating demo data",
    used = c(USED_SYNAPSE_ID),
    executed = c(USED_NB_GIT_URL)
)

store_to_synapse <- synStore(new_file_entity, activity = activity)
```


### c. Annotations
```{r}
# try retrieving annotation
synGetAnnotations(store_to_synapse$properties$id)
```

```{r}
#' set annotations
synSetAnnotations(
    store_to_synapse$properties$id, 
    list("annotatedBy" = "Aryton"))
```

## Part 3: Multi-File Uploads
This part will cover how to do bulk upload and bulk download to Synapse. It will use 'synapserutils' which is used for some utility use-cases built on top of 'synapser'
Documentation can be found here: https://docs.synapse.org/articles/uploading_in_bulk.html

### Bulk Upload
```{r}
synapserutils::syncToSynapse("manifest.tsv")
```

### Bulk Download
```{r}
synapserutils::syncFromSynapse("syn25257395")
```



