---
title: "Synapse Demo - Table Uploads"
author: "Aryton Tediarjo"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This R markdown is used for showcasing basic interaction with Synapse using RStudio. It will cover on how to utilize our Synapse R client to create table entity, which is a SQL query-able tables in Synapse interface. 

Full Documentation: https://r-docs.synapse.org/articles/tables.html


## Import Libraries
```{r warning=FALSE, message=FALSE}
library(data.table)
library(tidyverse)
library(synapser)
library(synapserutils)
synLogin()
```

## Global Variables
```{r}
SYNAPSE_PROJECT_ID <- "syn20816722" # where you will be saving all your files
FILEPATH_1 <- file.path("data", "sample_2.tsv")
FILEPATH_2 <- file.path("data", "sample_3.tsv")
TABLE_RESULTS <- "Synapse Demo Table"
```

```{r}
data <- fread(FILEPATH_1) %>% tibble::as_tibble()
data
```

### QnD storing table
A quick way to store table is to use available wrapper from Synapse R Client. It will infer data type of each column in the schema.
```{r}
#' wrapper create table function
table <- synapser::synBuildTable(
    TABLE_RESULTS, SYNAPSE_PROJECT_ID, data
)

#' store to synapse
table <- synStore(table)
```

### Custom Columns Types
A more manual way is to define each of the column schema individually.
```{r}
#' delete table
synapser::synDelete(table$tableId)

#' set custom columns
cols <- list(
    Column(name = "user", columnType = "STRING", maximumSize = 20),
    Column(name = "createdOn", columnType = "DATE"),
    Column(name = "x", columnType = "DOUBLE"),
    Column(name = "y", columnType = "DOUBLE"))

# set schema entity
schema <- Schema(name = TABLE_RESULTS, 
                 columns = cols, 
                 parent = SYNAPSE_PROJECT_ID)

#' recreate table entity
table <- Table(schema, data)
table <- synStore(table)
```

### Appending new rows to Synapse Tables
Appending new data can be achieved by storing a new data (given matching columns) to the previous table id.
```{r}
#' what is the table Id? 
tableId_in_synapse <- table$tableId

#' read data for appending
append_data <- fread(FILEPATH_2) %>% 
    tibble::as_tibble()

#' append to table
synStore(Table(tableId_in_synapse, append_data))
```


### Querying Data Examples
```{r}
cols <- c("data.json")
tbl_entity <- synTableQuery("SELECT * FROM syn25173598")
tbl_data <- tbl_entity$asDataFrame() %>%
  tidyr::pivot_longer(cols = all_of(cols), 
                      names_to = "column", 
                      values_to = "fileHandleId")

tbl_files <- synDownloadTableColumns(tbl_entity, columns = cols) %>% 
    tibble::enframe(.) %>%
    dplyr::select(fileHandleId = name, filePath = value)

tbl_data %>% 
  dplyr::inner_join(tbl_files, by = c("fileHandleId"))
```


### Extras: Useful code snippet to parse file attachment in Synapse Table
```{r}
#' Function to parse file attachment
#' @param synId the synapse table Id
#' @param cols the column of file attachment in table
parse_file_attachment <- function(synId, cols){
  tbl_entity <- synTableQuery(glue::glue("SELECT * FROM {synId}"))
  tbl_data <- tbl_entity$asDataFrame() %>%
    tidyr::pivot_longer(cols = all_of(cols), 
                        names_to = "column", 
                        values_to = "fileHandleId")

  tbl_files <- synDownloadTableColumns(tbl_entity, columns = cols) %>% 
    tibble::enframe(.) %>%
    dplyr::select(fileHandleId = name, filePath = value)

  tbl_data %>% 
    dplyr::inner_join(tbl_files, by = c("fileHandleId"))
}

parse_file_attachment(synId = "syn25173598", 
                      cols = c("data.json"))
```


